LME-MPM applied to quasi-brittle fracture.

### 2020-06-04 2020_EFM_MPM_Eigensoftening.tex
Mejora en  Edge-cracked square panel in mode I

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] / ([L] \cdot [T]^2)$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[M] / \cdot
[T]^2$. The absence of units is due to the current simulation will not be validated against experimental results, where
scale factor is relevant, but will be against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretization are required. On the one hand, a cartesian grid of
nodes is considered with a nodal spacing value of 0.025. On the other
hand, the plate will be modeled with a initial layout of four
particles per element and occupying the Gauss quadrature positions.\\

\PNA{NOMBRAR ALGO A LOS CASOS ANALITICOS O DE MICHAEL ORTIZ}

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both
loading and post-failure stages, when linear interpolation technique
is used. In contrast, \acrshort{lme} simulation does not exhibit these
spurious oscillations and remains linear until the failure. About the
strength, linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

>>>>

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current simulation will not be validated against experimental results, where
scale factor is relevant, but will be against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretization are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

\MMP{Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}
